---
title: "mtDNA Variants & GvHD Associations"
output: html_document
---

### ===============================

### (1) Setup & load data

### ===============================

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=TRUE, results='hide', warning=FALSE}
rm(list = ls())

## Load libraries
library(tidyverse)
library(seqMeta)
library(CompQuadForm)
library(twostageGWASsurvival)
library(MASS)
library(writexl)
library(readxl)
library(Matrix)
library(rmarkdown)
library(knitr)
library(data.table)
library(gtsummary)
library(dplyr)

# Set directory
setwd("")

```

#### 1-a) Prepare data for main analysis

```{r eval=TRUE, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))



# Check Multicolinearity
library(car)
library(performance)

vif_model <- lm(intxcgvhd ~ age+ donorage+
                                       prophx 
                                     + donorgp 
                                     +atgcampathgp_r 
                                     + pc1+ pc2+pc3+ pc4+pc5+ pc6+pc7+ pc8+pc9+ pc10, data =  pheno_fileV)
check_collinearity(vif_model)

vif(vif_model)'

```

#### Important points:

```         
1) The Seq Meta package is only available for R V4.0.2
2) You must define a values for repeated SNP position unless you will get same results for them, R will consider just one of them in dataset
3)  You must exclude SNPs with zero variation, additionally, SNPs with NA p-values obtained from singleSNPMeta results from both genomatrix and SNPinfo files, unless you will get error for SKAT-O and Burden test
4) Rename the patients ID in both files (pheno & genomatrix) with adding R to begining and replacing '.' with '_'

5) Define the PCs as numeric before ruining the analysis
6) All GVHD should be adjusted for age, dnrage, gvhprhrx_R1, donorgp, PC1, and PC2
7) You sould define snpInfo as data frame: snpInfo <- as.data.frame(snpInfo)
```

#### Input files:

```         
1) genomatrix: A matrix of mtDNA variants for each sample, where rows represent samples and columns represent variants. The values in the matrix indicate the presence or absence of specific variants in each sample. *Our File name: all494_2666_binary.vcf.txt*

2) SNPInfo: A data frame containing information about the mtDNA variants, including their names, positions, and other relevant annotations. This file is used to provide context for the variants being analyzed in the SKAT-O test. *Our File name: SNP_POS_ALT_Func_2666_494_R2.xlsx*

3) Pheno_file: Clinical/Demographic information of samples. *Our Final File name: MDS_imputed_Clinical_MDS&donor_March9th_2026.xlsx*
```

##### 1-b) Import the data

```{r load-data, eval=TRUE, message=FALSE, include=TRUE, paged.print=TRUE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Recipient data
pheno_file = read_excel('MDS_imputed_Clinical_MDS&donor_March14th_2026.xlsx')
genomatrix = read.table("all494_2666_binary.vcf.txt",header=TRUE)
snpInfo = read_excel('SNP_POS_ALT_Func_2666_494_R2.xlsx')

# Define variables types
pheno_file <- pheno_file %>%
  mutate(
    # Numeric variables
    across(
      c(
        event_aGvhD24, event_aGvhD34,
        intxagvhd24, intxagvhd34, intxcgvhd, cgvhd,
        PC1, PC2, dnrage, age
      ),
      ~ as.numeric(as.character(.x))
    ),
    
    # Factor variables
    across(
      c(gvhprhrx_R1, donorgp),
      as.factor
    )
  )

# Donor data
pheno_fileD = read_excel('MDS_imputed_Clinical_MDS&donor_March14th_2026.xlsx')
genomatrixD = read.table("all494_1825_binary.vcf.txt",header=TRUE)
snpInfoD = read_excel('SNP_POS_ALT_Func_1825_494_R2.xlsx')

# Define variables types
pheno_fileV <- pheno_fileV %>%
  mutate(
    # Numeric variables
    across(
      c(event_aGvhD24, event_aGvhD34,
        intxagvhd24, intxagvhd34, intxcgvhd, cgvhd,
        pc1, pc2, donorage, age),
      as.numeric
    ),
    
    # Factor variables
    across(
      c(donorgp, prophx),
      as.factor
    )
  )




# Read validation dataset- NCI
genomatrixV = read.table("MDS295_2942_binary.txt",header=TRUE)
snpInfoV = read.csv('SNV_info_excluded_non-coding.csv',header=TRUE)
pheno_fileV = read_excel("MDS295_clinical_R4.xlsx")

pheno_fileV$event_aGvhD24[pheno_fileV$event_aGvhD24==98] <- NA
pheno_fileV$event_aGvhD34[pheno_fileV$event_aGvhD34==98] <- NA
pheno_fileV$cgvhd[pheno_fileV$cgvhd== 99] <- NA

# Define variables types
pheno_fileV <- pheno_fileV %>%
  mutate(
    across(
      c(
        event_aGvhD24, event_aGvhD34,
        intxagvhd24, intxagvhd34, intxcgvhd, cgvhd,
        pc1, pc2, donorage, age
      ),
      ~ as.numeric(as.character(.x))
    ),
    
    across(
      c(donorgp, prophx),
      as.factor
    )
  )


# There is no Remove rows where prophx and cgvhd is NA

```

##### 1-c) Create genomatrix for cox model to merge snpInfo and genomatrix info- recipients

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=TRUE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))


#//////////#
I = match(snpInfo$POS, genomatrix$POS)

genomatrix = genomatrix[I,]

genomatrix = t(genomatrix[,-c(1:3)])

colnames(genomatrix) = snpInfo$SNP


samples = row.names(genomatrix)
I = match(as.character(pheno_file$formattedRID_LabCorpID),samples)
genomatrix = genomatrix[I,]

rownames(genomatrix) <- pheno_file$formattedRID_LabCorpID


# You can Save genomatrix as csv file
# write.csv(as.data.frame(genomatrix), "genomatrix.csv", row.names=TRUE)


# Check the exact name in dataset (snpInfo) maybe it is Gene18
aggregate_by = 'gene18'


# Define snpInfo as data frame, I got error due to this took long time to figure out 
snpInfo <- as.data.frame(snpInfo)


#//////////////////////////////////////////////////////////////////////////////#
#//////////////////////////////////////////////////////////////////////////////#
# 2) Define repeated position in genomatrix and SNPinfo files in donors

#//////////#
ID = match(snpInfoD$POS, genomatrixD$POS)

genomatrixD = genomatrixD[ID,]

genomatrixD = t(genomatrixD[,-c(1:3)])

colnames(genomatrixD) = snpInfoD$SNP


samplesD = row.names(genomatrixD)
ID = match(as.character(pheno_fileD$Donor_formattedRID_LabCorpID),samplesD)
genomatrixD = genomatrixD[ID,]

rownames(genomatrixD) <- pheno_fileD$Donor_formattedRID_LabCorpID


# Check the exact name in dataset maybe it is Gene18
aggregate_byD = 'Gene18'


# Define snpInfo as data frame
snpInfoD <- as.data.frame(snpInfoD)



#//////////////////////////////////////////////////////////////////////////////#
#//////////////////////////////////////////////////////////////////////////////#
# 3) Define repeated position in genomatrixV and SNPinfo files in validation
IV = match(snpInfoV$POS, genomatrixV$POS)

genomatrixV = genomatrixV[IV,]

genomatrixV = t(genomatrixV[,-c(1:3)])

colnames(genomatrixV) = snpInfoV$SNP


samplesV = row.names(genomatrixV)
IV = match(as.character(pheno_fileV$subject_id),samplesV)
genomatrixV = genomatrixV[IV,]

rownames(genomatrixV) <- pheno_fileV$subject_id



aggregate_byV = 'Gene18b'

```

## Recipients mtDNA & GvHD  (2-4)

### ===============================

### (2) Associations between mtDNA & aGvHD24 - Recipients

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD24 in MDS patients (Recipients)

##### 2-a) Run the cause-specific Cox SKAT-O model for aGvHD24 in recipients to find variants with NA P-values

```         
Before running the SKAT-O analysis, first check for any NA p‑values in the single‑SNV analysis. Based on those results, we need to remove variants with zero variation or NA p‑values; otherwise, SKAT‑O and the burden test will return errors
```

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

#----------------------------#
# Check Cox Model assumptions
cox_base_aGVHD24 <- coxph(
  Surv(intxagvhd24, event_aGvhD24) ~
    age + dnrage + gvhprhrx_R1 + donorgp + PC1 + PC2,
  data = pheno_file,
  x = TRUE
)

PH_base_aGVHD24 <- cox.zph(cox_base_aGVHD24)
PH_base_aGVHD24
PH_base_aGVHD24$table

plot(PH_base_aGVHD24)
#----------------------------#

# Main cox model
Cox_model_aGVHD24_MDS<-prepCox(genomatrix, Surv(intxagvhd24,event_aGvhD24) ~ age+ dnrage+  gvhprhrx_R1+ donorgp  + PC1+PC2,SNPInfo = snpInfo, aggregateBy = aggregate_by, data = pheno_file ,verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD24_MDS<-singlesnpMeta(Cox_model_aGVHD24_MDS, SNPInfo = snpInfo, aggregateBy = aggregate_by)

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD24_MDS_NA <- singleSNPMeta_aGVHD24_MDS %>%
  filter(is.na(p))

singleSNPMeta_aGVHD24_MDS_NA <- singleSNPMeta_aGVHD24_MDS[is.na(singleSNPMeta_aGVHD24_MDS$p), ]

print(singleSNPMeta_aGVHD24_MDS_NA)

# Save SNV results
write.table(singleSNPMeta_aGVHD24_MDS, file = "Results_singleSNPMeta_aGvHD24_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

##### 2-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again
```

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}

# 1. Filter SNPs


# 2. Run Cox model
Fit_Cox_model_aGVHD24_MDS <- prepCox(genomatrix,Surv(intxagvhd24, event_aGvhD24) ~ age + dnrage + gvhprhrx_R1 +donorgp+ PC1 + PC2,SNPInfo = snpInfo, aggregateBy = aggregate_by,  data = pheno_file,verbose = TRUE)

# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_aGVHD24_MDS <- skatOMeta(Fit_Cox_model_aGVHD24_MDS, 
                                         SNPInfo = snpInfo, 
                                         aggregateBy = aggregate_by, 
                                         verbose = TRUE)
# 4. Run Burden test
Fit_burden_Cox_aGVHD24_MDS <- burdenMeta(Fit_Cox_model_aGVHD24_MDS, 
                                         SNPInfo = snpInfo, 
                                         aggregateBy = aggregate_by)

# 4. Save outputs 
write.table(Fit_Cox_skato_model_aGVHD24_MDS, file = "Results_Cox_skato_aGVHD24_MDS_March2026b.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD24_MDS, file = "Results_Cox_burden_aGVHD24_MDS_March2026b.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD24_MDS, "Results_aGVHD24_skato_MDSb.rds")
saveRDS(Fit_burden_Cox_aGVHD24_MDS, "Results_aGVHD24_burden_MDSb.rds")
```

#### 2-c) Results of SKAT-O - aGvHD24 in recipients

```{r}

print(Fit_Cox_skato_model_aGVHD24_MDS)
```

#### 2-d) Results of Burden test - aGvHD24 in recipients

```{r}
print(Fit_burden_Cox_aGVHD24_MDS)
```

### ===============================

### (3) Associations between mtDNA & aGvHD34 - Recipients

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD34 in recipients

##### 3-a) Run the cause-specific Cox SKAT-O model for aGvHD34 in recipients to check anye NA p‑values in the single‑SNV analysis

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

#----------------------------#
# Check Cox Model assumptions: aGvHD34
cox_base_aGVHD34 <- coxph(
  Surv(intxagvhd34, event_aGvhD34) ~
    age + dnrage + gvhprhrx_R1 + donorgp + PC1 + PC2,
  data = pheno_file,
  x = TRUE
)

PH_base_aGVHD34 <- cox.zph(cox_base_aGVHD34)
PH_base_aGVHD34$table

plot(PH_base_aGVHD34)
#----------------------------#

# Main cox model
Cox_model_aGVHD34_MDS<-prepCox(genomatrix, Surv(intxagvhd34,event_aGvhD34) ~ age+ dnrage+  gvhprhrx_R1+ donorgp + PC1+PC2,SNPInfo = snpInfo, aggregateBy = aggregate_by, data = pheno_file ,verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD34_MDS<-singlesnpMeta(Cox_model_aGVHD34_MDS, SNPInfo = snpInfo, aggregateBy = aggregate_by)

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD34_MDS_NA <- singleSNPMeta_aGVHD34_MDS %>%
  filter(is.na(p))

singleSNPMeta_aGVHD34_MDS_NA <- singleSNPMeta_aGVHD34_MDS[is.na(singleSNPMeta_aGVHD34_MDS$p), ]

print(singleSNPMeta_aGVHD34_MDS_NA)

# Save SNV results
write.table(singleSNPMeta_aGVHD34_MDS, file = "Results_singleSNPMeta_aGvHD34_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again if there is any
The p-values with NA = 0 
```

##### 3-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))


# 1. Run Cox model
Fit_Cox_model_aGVHD34_MDS <- prepCox(genomatrix,Surv(intxagvhd34, event_aGvhD34) ~ age+ dnrage+  gvhprhrx_R1+ donorgp+ PC1+PC2,SNPInfo = snpInfo,aggregateBy = aggregate_by, data = pheno_file,verbose = TRUE)
str(pheno_file)
# 2. Run Cox  Skat-o model
Fit_Cox_skato_model_aGVHD34_MDS <- skatOMeta(Fit_Cox_model_aGVHD34_MDS, 
                                         SNPInfo = snpInfo, 
                                         aggregateBy = aggregate_by, 
                                         verbose = TRUE)
# 3. Run Burden test 
Fit_burden_Cox_aGVHD34_MDS <- burdenMeta(Fit_Cox_model_aGVHD34_MDS, 
                                         SNPInfo = snpInfo, 
                                         aggregateBy = aggregate_by)

# 4. Save outputs
write.table(Fit_Cox_skato_model_aGVHD34_MDS, file = "Results_Cox_skato_aGVHD34_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD34_MDS, file = "Results_Cox_burden_aGVHD34_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD34_MDS, "results_agvhd34_skato_MDS.rds")
saveRDS(Fit_burden_Cox_aGVHD34_MDS, "results_agvhd34_burden_MDS.rds")
```

#### 3-c) Results of SKAT-O - aGvHD34 in recipients

```{r}
print(Fit_Cox_skato_model_aGVHD34_MDS)
```

#### 3-d) Results of Burden test - aGvHD34 in recipients

```{r}
print(Fit_burden_Cox_aGVHD34_MDS)
```

### ===============================

### (4) Associations between mtDNA & cGvHD - Recipients

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and chronic GVHD in recipients

##### 4-a) Run the cause-specific Cox SKAT-O model for cGvHD in recipients to check anye NA p‑values in the single‑SNV analysis

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

#----------------------------#
# Check Cox Model assumptions: aGvHD34
cox_base_cGVHD <- coxph(
  Surv(intxcgvhd, cgvhd) ~
    age + dnrage + gvhprhrx_R1 + donorgp + PC1 + PC2,
  data = pheno_file,
  x = TRUE
)

PH_base_cGVHD <- cox.zph(cox_base_cGVHD)
PH_base_cGVHD$table

plot(PH_base_cGVHD)
# all are >0.1 and plots dont show any violation of PH assumption
#----------------------------#


# Main cox model
Cox_model_cGvHD_MDS<-prepCox(genomatrix, Surv(intxcgvhd,cgvhd) ~ age+ dnrage+  gvhprhrx_R1+ donorgp  + PC1+PC2,SNPInfo = snpInfo, aggregateBy = aggregate_by, data = pheno_file ,verbose = TRUE)


# Single SNV
singleSNPMeta_cGvHD_MDS<-singlesnpMeta(Cox_model_cGvHD_MDS, SNPInfo = snpInfo, aggregateBy = aggregate_by)

# Print SNVs with P-value = NA
singleSNPMeta_cGvHD_MDS_NA <- singleSNPMeta_cGvHD_MDS %>%
  filter(is.na(p))

singleSNPMeta_cGvHD_MDS_NA <- singleSNPMeta_cGvHD_MDS[is.na(singleSNPMeta_cGvHD_MDS$p), ]

print(singleSNPMeta_cGvHD_MDS_NA)

# Save SNV results
write.table(singleSNPMeta_cGvHD_MDS, file = "Results_singleSNPMeta_cGvHD_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again if there is any
The p-values with NA = 0 
```

##### 4-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}

# 1. Filter SNPs: We didn't observed any NA for P-valuse in single variant analysis

SNP_to_removec <- c('snp64',
'snp237',
'snp280',
'snp420',
'snp434',
'snp546',
'snp718',
'snp865',
'snp913',
'snp1261',
'snp1302',
'snp1740',
'snp1958',
'snp2160')
snpInfo_cGvHD <- snpInfo %>% filter(!Name %in% SNP_to_removec)

# 2. Run Cox model
Fit_Cox_model_cGvHD_MDS <- prepCox(genomatrix,Surv(intxcgvhd,cgvhd) ~ age+ dnrage+  gvhprhrx_R1+ donorgp  + PC1+PC2,SNPInfo = snpInfo_cGvHD,aggregateBy = aggregate_by, data = pheno_file, verbose = TRUE)

# is any NA for intxcgvhd
pheno_file$intxcgvhd[is.na(pheno_file$intxcgvhd)]


# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_cGvHD_MDS <- skatOMeta(Fit_Cox_model_cGvHD_MDS, 
                                         SNPInfo = snpInfo_cGvHD, 
                                         aggregateBy = aggregate_by, 
                                         verbose = TRUE)
# 4. Run Burden test 
Fit_burden_Cox_cGvHD_MDS <- burdenMeta(Fit_Cox_model_cGvHD_MDS, 
                                         SNPInfo = snpInfo_cGvHD, 
                                         aggregateBy = aggregate_by)

# 5. Save outputs
write.table(Fit_Cox_skato_model_cGvHD_MDS, file = "Results_Cox_skato_cGvHD_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_cGvHD_MDS, file = "Results_Cox_burden_cGvHD_MDS_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_cGvHD_MDS, "results_cGvHD_skato_MDS.rds")
saveRDS(Fit_burden_Cox_cGvHD_MDS, "results_cGvHD_burden_MDS.rds")
```

#### 4-c) Results of SKAT-O - cGvHD in recipients

```{r}
print(Fit_Cox_skato_model_cGvHD_MDS)
```

#### 4-d) Results of Burden test - cGvHD in recipients

```{r}
print(Fit_burden_Cox_cGvHD_MDS)
```

## Donor mtDNA & GvHD (5-7)

### ===============================

### (5) Associations between mtDNA & aGvHD24 - donors

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD24 in donors

##### 5-a) Run the cause-specific Cox SKAT-O model for aGvHD24 in donors to find variants with NA P-values

```         
Before running the SKAT-O analysis, first check for any NA p‑values in the single‑SNV analysis. Based on those results, we need to remove variants with zero variation or NA p‑values; otherwise, SKAT‑O and the burden test will return errors
```

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Main cox model
Cox_model_aGVHD24_Donor<-prepCox(genomatrixD, Surv(intxagvhd24,event_aGvhD24) ~ age+ dnrage+  gvhprhrx_R1+ donorgp +pcD1+pcD2,SNPInfo = snpInfoD, aggregateBy = aggregate_byD, data = pheno_fileD ,verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD24_Donor<-singlesnpMeta(Cox_model_aGVHD24_Donor, SNPInfo = snpInfoD, aggregateBy = aggregate_byD)

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD24_Donor_NA <- singleSNPMeta_aGVHD24_Donor %>%
  filter(is.na(p))

singleSNPMeta_aGVHD24_Donor_NA <- singleSNPMeta_aGVHD24_Donor[is.na(singleSNPMeta_aGVHD24_Donor$p), ]

print(singleSNPMeta_aGVHD24_Donor_NA)

# Save SNV results
write.table(singleSNPMeta_aGVHD24_Donor, file = "singleSNPMeta_aGvHD24_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

##### 5-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again
```

```{r  eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}

# 1. Filter aGVHD24 Donor SNPs

# 2. Run Cox model aGVHD24 Donor
Fit_Cox_model_aGVHD24_Donor <- prepCox(genomatrixD,Surv(intxagvhd24, event_aGvhD24) ~ age + dnrage + gvhprhrx_R1 +donorgp+pcD1+pcD2, SNPInfo = snpInfoD,aggregateBy = aggregate_byD, 
                                     data = pheno_fileD, 
                                     verbose = TRUE)

# 3. Run Cox  Skat-o model aGVHD24 Donor
Fit_Cox_skato_model_aGVHD24_Donor <- skatOMeta(Fit_Cox_model_aGVHD24_Donor, 
                                         SNPInfo = snpInfoD, 
                                         aggregateBy = aggregate_byD, 
                                         verbose = TRUE)

# 4. Run Burden test aGVHD24 Donor
Fit_burden_Cox_aGVHD24_Donor <- burdenMeta(Fit_Cox_model_aGVHD24_Donor, 
                                         SNPInfo =  snpInfoD, 
                                         aggregateBy = aggregate_byD)

# 4. Save outputs 
write.table(Fit_Cox_skato_model_aGVHD24_Donor, file = "Results_Cox_skato_aGVHD24_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD24_Donor, file = "Results_Cox_burden_aGVHD24_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD24_Donor, "results_aGVHD24_skato_Donor.rds")
saveRDS(Fit_burden_Cox_aGVHD24_Donor, "results_aGVHD24_burden_Donor.rds")
```

#### 5-c) Results of SKAT-O - aGvHD24 in donors

```{r}

print(Fit_Cox_skato_model_aGVHD24_Donor)
```

#### 5-d) Results of Burden test - aGvHD24 in donors

```{r}
print(Fit_burden_Cox_aGVHD24_Donor)
```

### ===============================

### (6) Associations between mtDNA & aGvHD34 - donors

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD34 in donors

##### 6-a) Run the cause-specific Cox SKAT-O model for aGvHD34 in donors to check anye NA p‑values in the single‑SNV analysis

```{r donor-agvhd34-cox, eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Main cox model
Cox_model_aGVHD34_Donor<-prepCox(genomatrixD, Surv(intxagvhd34,event_aGvhD34) ~ age+ dnrage+  gvhprhrx_R1+ donorgp + pcD1+pcD2,SNPInfo = snpInfoD, aggregateBy = aggregate_byD, data = pheno_fileD ,verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD34_Donor<-singlesnpMeta(Cox_model_aGVHD34_Donor, SNPInfo = snpInfoD, aggregateBy = aggregate_byD)

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD34_Donor_NA <- singleSNPMeta_aGVHD34_Donor %>%
  filter(is.na(p))

singleSNPMeta_aGVHD34_Donor_NA <- singleSNPMeta_aGVHD34_Donor[is.na(singleSNPMeta_aGVHD34_Donor$p), ]

print(singleSNPMeta_aGVHD34_Donor_NA)

# Save SNV results
write.table(singleSNPMeta_aGVHD34_Donor, file = "Results_singleSNPMeta_aGvHD34_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again if there is any
The p-values with NA = 0: same as  aGvHD24 "snp302"
```

##### 6-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# 1. Filter SNPs: exclude NA for P-valuse in single variant analysis
SNP_to_removeD <- c("snp302")
snpInfoD_aGvHD34 <- snpInfoD %>% filter(!Name %in% SNP_to_removeD)




# 2. Run Cox model
Fit_Cox_model_aGVHD34_Donor <- prepCox(genomatrixD,Surv(intxagvhd34, event_aGvhD34) ~ age+ dnrage+ gvhprhrx_R1+ donorgp  + pcD1+pcD2,SNPInfo = snpInfoD_aGvHD34, 
                                     aggregateBy = aggregate_byD, 
                                     data = pheno_fileD, 
                                     verbose = TRUE)


# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_aGVHD34_Donor <- skatOMeta(Fit_Cox_model_aGVHD34_Donor, 
                                         SNPInfo = snpInfoD_aGvHD34, 
                                         aggregateBy = aggregate_byD, 
                                         verbose = TRUE)
# 4. Run Burden test 
Fit_burden_Cox_aGVHD34_Donor <- burdenMeta(Fit_Cox_model_aGVHD34_Donor, 
                                         SNPInfo = snpInfoD_aGvHD34, 
                                         aggregateBy = aggregate_byD)

# 5. Save outputs
write.table(Fit_Cox_skato_model_aGVHD34_Donor, file = "Results_Cox_skato_aGVHD34_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD34_Donor, file = "Results_Cox_burden_aGVHD34_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD34_Donor, "results_aGVHD34_skato_Donor.rds")
saveRDS(Fit_burden_Cox_aGVHD34_Donor, "results_aGVHD34_burden_Donor.rds")
```

#### 6-c) Results of SKAT-O - aGvHD34 in donors

```{r}
print(Fit_Cox_skato_model_aGVHD34_Donor)
```

#### 6-d) Results of Burden test - aGvHD34 in donors

```{r}
print(Fit_burden_Cox_aGVHD34_Donor)
```

### ===============================

### (7) Associations between mtDNA & cGvHD - donors

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and chronic GVHD in donors

##### 7-a) Run the cause-specific Cox SKAT-O model for cGvHD in donors to check anye NA p‑values in the single‑SNV analysis

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))


# Main cox model
Cox_model_cGvHD_Donor<-prepCox(genomatrixD, Surv(intxcgvhd,cgvhd) ~ age+ dnrage+  gvhprhrx_R1+ donorgp  + pcD1+pcD2,SNPInfo = snpInfoD, aggregateBy = aggregate_byD, data = pheno_fileD ,verbose = TRUE)


# Single SNV
singleSNPMeta_cGvHD_Donor<-singlesnpMeta(Cox_model_cGvHD_Donor, SNPInfo = snpInfoD, aggregateBy = aggregate_byD)

# Print SNVs with P-value = NA
singleSNPMeta_cGvHD_Donor_NA <- singleSNPMeta_cGvHD_Donor %>%
  filter(is.na(p))

singleSNPMeta_cGvHD_Donor_NA <- singleSNPMeta_cGvHD_Donor[is.na(singleSNPMeta_cGvHD_Donor$p), ]

print(singleSNPMeta_cGvHD_Donor_NA)

# Save SNV results
write.table(singleSNPMeta_cGvHD_Donor, file = "Results_singleSNPMeta_cGvHD_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

##### 7-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}

# 1. Filter SNPs: observed  NA for P-valuse in single variant analysis
SNP_to_removeDc <- c('snp601',	'snp675',	'snp690',	'snp706',	'snp708',	'snp716',	'snp799',	'snp944',	'snp38',	'snp94',	'snp1734',	'snp1488',	'snp1507',	'snp1582',	'snp493',	'snp519',	'snp542',	'snp1014',	'snp1135',	'snp1060',	'snp1265',	'snp1418',	'snp1444',	'snp245',	'snp290',	'snp302',	'snp184',	'snp1204',	'snp1624'
)
snpInfoD_cGvHD <- snpInfoD %>% filter(!Name %in% SNP_to_removeDc)

# 2. Run Cox model
Fit_Cox_model_cGvHD_Donor <- prepCox(genomatrixD, 
                                     Surv(intxcgvhd,cgvhd) ~ age+ dnrage+  gvhprhrx_R1+ donorgp + pcD1+pcD2,
                                     SNPInfo =  snpInfoD_cGvHD, 
                                     aggregateBy = aggregate_byD, 
                                     data = pheno_fileD, 
                                     verbose = TRUE)
# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_cGvHD_Donor <- skatOMeta(Fit_Cox_model_cGvHD_Donor, 
                                         SNPInfo = snpInfoD_cGvHD, 
                                         aggregateBy = aggregate_byD, 
                                         verbose = TRUE)
# 4. Run Burden test 
Fit_burden_Cox_cGvHD_Donor <- burdenMeta(Fit_Cox_model_cGvHD_Donor, 
                                         SNPInfo = snpInfoD_cGvHD, 
                                         aggregateBy = aggregate_byD)

# 5. Save outputs
write.table(Fit_Cox_skato_model_cGvHD_Donor, file = "Results_Cox_skato_cGvHD_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_cGvHD_Donor, file = "Results_Cox_burden_cGvHD_Donor_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_cGvHD_Donor, "results_cGVHD_skato_Donor.rds")
saveRDS(Fit_burden_Cox_cGvHD_Donor, "results_cGVHD_burden_Donor.rds")
```

#### 7-c) Results of SKAT-O - cGvHD in donors

```{r}
print(Fit_Cox_skato_model_cGvHD_Donor)
```

#### 7-d) Results of Burden test - cGvHD in donors

```{r}
print(Fit_burden_Cox_cGvHD_Donor)
```

## Validation-recipients mtDNA & GvHD (8-10)

### ===============================

### (8) Associations between mtDNA & aGvHD24 - Recipients validation

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD24 in MDS patients (validation)

##### 8-a) Run the cause-specific Cox SKAT-O model for aGvHD24 in recipients to find variants with NA P-values

```         
Before running the SKAT-O analysis, first check for any NA p‑values in the single‑SNV analysis. Based on those results, we need to remove variants with zero variation or NA p‑values; otherwise, SKAT‑O and the burden test will return errors
```

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Main cox model
Cox_model_aGVHD24_MDS_validation<-prepCox(genomatrixV, Surv(day_aGvHD24,event_aGvhD24) ~ age+ donorage+  prophx+ donorgp + pc1+ pc2,SNPInfo = snpInfoV, snpNames="SNP",aggregateBy = aggregate_byV, data = pheno_fileV ,verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD24_MDS_validation<-singlesnpMeta(Cox_model_aGVHD24_MDS_validation,snpNames="SNP", SNPInfo = snpInfoV, aggregateBy = aggregate_byV)

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD24_MDS_NA_validation <- singleSNPMeta_aGVHD24_MDS_validation %>%
  filter(is.na(p))

singleSNPMeta_aGVHD24_MDS_NA_validation <- singleSNPMeta_aGVHD24_MDS_validation[is.na(singleSNPMeta_aGVHD24_MDS_validation$p), ]

print(singleSNPMeta_aGVHD24_MDS_NA_validation)

# Save SNV results
write.table(singleSNPMeta_aGVHD24_MDS_validation, file = "Results_singleSNPMeta_aGvHD24_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

##### 8-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again
```

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}

# 1. Run Cox model
Fit_Cox_model_aGVHD24_MDS_validation <- prepCox(genomatrixV,Surv(day_aGvHD24, event_aGvhD24) ~ age+ donorage+  prophx+ donorgp + pc1+ pc2,SNPInfo = snpInfoV, snpNames="SNP", aggregateBy = aggregate_byV, data = pheno_fileV,verbose = TRUE)

# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_aGVHD24_MDS_validation <- skatOMeta(Fit_Cox_model_aGVHD24_MDS_validation, 
                                         SNPInfo = snpInfoV, 
                                         aggregateBy = aggregate_byV, snpNames="SNP",
                                         verbose = TRUE)
# 4. Run Burden test
Fit_burden_Cox_aGVHD24_MDS_validation <- burdenMeta(Fit_Cox_model_aGVHD24_MDS_validation, 
                                         SNPInfo = snpInfoV,snpNames="SNP", 
                                         aggregateBy = aggregate_byV)

# 4. Save outputs 
write.table(Fit_Cox_skato_model_aGVHD24_MDS_validation, file = "Results_Cox_skato_aGVHD24_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD24_MDS_validation, file = "Results_Cox_burden_aGVHD24_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD24_MDS_validation, "Results_aGVHD24_skato_validation_MDS.rds")
saveRDS(Fit_burden_Cox_aGVHD24_MDS_validation, "Results_aGVHD24_burden_validation_MDS.rds")
```

#### 8-c) Results of SKAT-O - aGvHD24 in recipients- validation

```{r}

print(Fit_Cox_skato_model_aGVHD24_MDS_validation)
```

#### 8-d) Results of Burden test - aGvHD24 in recipients- validation

```{r}
print(Fit_burden_Cox_aGVHD24_MDS_validation)
```

### ===============================

### (9) Associations between mtDNA & aGvHD34 - Recipients validation

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and acute GVHD34 in recipients

##### 9-a) Run the cause-specific Cox SKAT-O model for aGvHD34 in recipients to check anye NA p‑values in the single‑SNV analysis

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Main cox model
Cox_model_aGVHD34_MDS_validation<-prepCox(genomatrixV, Surv(day_aGvHD34,event_aGvhD34) ~ age+ donorage+  prophx+ donorgp+ pc1+ pc2,SNPInfo = snpInfoV, aggregateBy = aggregate_byV, data = pheno_fileV ,snpNames="SNP",verbose = TRUE)


# Single SNV
singleSNPMeta_aGVHD34_MDS_validation<-singlesnpMeta(Cox_model_aGVHD34_MDS_validation, SNPInfo = snpInfoV, aggregateBy = aggregate_byV,snpNames="SNP")

# Print SNVs with P-value = NA
singleSNPMeta_aGVHD34_MDS_NA_validation <- singleSNPMeta_aGVHD34_MDS_validation %>%
  filter(is.na(p))

singleSNPMeta_aGVHD34_MDS_NA_validation <- singleSNPMeta_aGVHD34_MDS_validation[is.na(singleSNPMeta_aGVHD34_MDS_validation$p), ]

print(singleSNPMeta_aGVHD34_MDS_NA_validation)

# Save SNV results
write.table(singleSNPMeta_aGVHD34_MDS_validation, file = "Results_singleSNPMeta_aGvHD34_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again if there is any
The p-values with NA = 0 
```

##### 9-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# 1. Filter SNPs: We didn't observed any NA for P-valuse in single variant analysis


# 2. Run Cox model
Fit_Cox_model_aGVHD34_MDS_validation <- prepCox(genomatrixV,Surv(day_aGvHD34, event_aGvhD34) ~ age + donorage+ prophx+ donorgp+ pc1+ pc2, SNPInfo = snpInfoV, 
                                     aggregateBy = aggregate_byV, 
                                     data = pheno_fileV, snpNames="SNP",
                                     verbose = TRUE)
# 3. Run Cox  Skat-o model
Fit_Cox_skato_model_aGVHD34_MDS_validation <- skatOMeta(Fit_Cox_model_aGVHD34_MDS_validation, 
                                         SNPInfo = snpInfoV, 
                                         aggregateBy = aggregate_byV,snpNames="SNP", 
                                         verbose = TRUE)
# 4. Run Burden test 
Fit_burden_Cox_aGVHD34_MDS_validation <- burdenMeta(Fit_Cox_model_aGVHD34_MDS_validation, 
                                         SNPInfo = snpInfoV,snpNames="SNP", 
                                         aggregateBy = aggregate_byV)

# 5. Save outputs
write.table(Fit_Cox_skato_model_aGVHD34_MDS_validation, file = "Results_Cox_skato_aGVHD34_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_aGVHD34_MDS_validation, file = "Results_Cox_burden_aGVHD34_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_aGVHD34_MDS_validation, "results_agvhd34_skato_MDS_validation.rds")
saveRDS(Fit_burden_Cox_aGVHD34_MDS_validation, "results_agvhd34_burden_MDS_validation.rds")
```

#### 9-c) Results of SKAT-O - aGvHD34 in recipients- validation

```{r}
print(Fit_Cox_skato_model_aGVHD34_MDS_validation)
```

#### 9-d) Results of Burden test - aGvHD34 in recipients- validation

```{r}
print(Fit_burden_Cox_aGVHD34_MDS_validation)
```


### ===============================

### (10) Associations between mtDNA & cGvHD - Recipients validation

### ===============================

### Aim: Cause-specific Cox SKAT-O to test the association between mtDNA variants and chronic GVHD in recipients

##### 10-a) Run the cause-specific Cox SKAT-O model for cGvHD in recipients to check anye NA p‑values in the single‑SNV analysis

```{r eval=TRUE, message=FALSE, include=TRUE, paged.print=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(engine.opts = list(echo = FALSE, warning=FALSE,
                                         perl = '-Mstrict -Mwarnings',
                                         bash = '-o errexit'
))

# Main cox model
Cox_model_cGvHD_MDS_validation<-prepCox(genomatrixV, Surv(intxcgvhd,cgvhd) ~ age+ donorage+  prophx+ donorgp + pc1+ pc2, SNPInfo = snpInfoV, aggregateBy = aggregate_byV, data = pheno_fileV ,snpNames="SNP",verbose = TRUE)


# Single SNV
singleSNPMeta_cGvHD_MDS_validation<-singlesnpMeta(Cox_model_cGvHD_MDS_validation, SNPInfo = snpInfoV, snpNames="SNP",aggregateBy = aggregate_byV)

# Print SNVs with P-value = NA
singleSNPMeta_cGvHD_MDS_NA_validation <- singleSNPMeta_cGvHD_MDS_validation %>%
  filter(is.na(p))

singleSNPMeta_cGvHD_MDS_NA_validation <- singleSNPMeta_cGvHD_MDS_validation[is.na(singleSNPMeta_cGvHD_MDS_validation$p), ]

print(singleSNPMeta_cGvHD_MDS_NA_validation)

# Save SNV results
write.table(singleSNPMeta_cGvHD_MDS_validation, file = "Results_singleSNPMeta_cGvHD_MDS__validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```         
We need to remove the variants with NA P-values from SNPINFO file, then run the Cox SKAT-O model again if there is any
The p-values with NA = 0 
```

##### 10-b) Run the Cox SKAT-O model again after removing variants with NA P-values

```{r eval=TRUE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE, results='hide'}

# 1. Run Cox model
Fit_Cox_model_cGvHD_MDS_validation <- prepCox(genomatrixV, 
                                     Surv(intxcgvhd,cgvhd) ~ age+ donorage+prophx 
                                     + donorgp
                                     + pc1+ pc2
                                     ,SNPInfo = snpInfoV ,snpNames="SNP",
                                     aggregateBy = aggregate_byV, 
                                     data = pheno_fileV,
                                     verbose = TRUE)
# 2. Run Cox  Skat-o model
Fit_Cox_skato_model_cGvHD_MDS_validation <- skatOMeta(Fit_Cox_model_cGvHD_MDS_validation, 
                                         SNPInfo = snpInfoV , 
                                         aggregateBy = aggregate_byV,snpNames="SNP", 
                                         verbose = TRUE)


# 3. Run Burden test 
Fit_burden_Cox_cGvHD_MDS_validation <- burdenMeta(Fit_Cox_model_cGvHD_MDS_validation, 
                                         SNPInfo = snpInfoV, snpNames="SNP" ,
                                         aggregateBy = aggregate_byV)

# 4. Save outputs
write.table(Fit_Cox_skato_model_cGvHD_MDS_validation, file = "Results_Cox_skato_cGvHD_MDS_validation_March2026_without_only_age&drgroup_depliation.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(Fit_burden_Cox_cGvHD_MDS_validation, file = "Results_Cox_burden_cGvHD_MDS_validation_March2026.txt", sep = "\t", row.names = FALSE, quote = FALSE)

saveRDS(Fit_Cox_skato_model_cGvHD_MDS_validation, "results_cGvHD_skato_MDS_validation.rds")
saveRDS(Fit_burden_Cox_cGvHD_MDS_validation, "results_cGvHD_burden_MDS_validation.rds")
```

#### 10-c) Results of SKAT-O - cGvHD in recipients- validation

```{r}
print(Fit_Cox_skato_model_cGvHD_MDS_validation)
```

#### 10-d) Results of Burden test - cGvHD in recipients- validation

```{r}
print(Fit_burden_Cox_cGvHD_MDS_validation)
```
